Discussion

### 2020-08-16 text/discussion.Rmd

Finally, we must remark on the limitations of our work exemplified in two of the
classes, 1 and 10, for which we did not find many distinctive associations with
pathways (**figure 5B, C**). Class 1 had a relatively low proportion of mutated
pathways across the board, although it is broadly enriched in many of the same
tumor-promoting pathways common to the other groups. This likely does not
reflect a true difference in stages of tumor evolution, as our data show clearly
that this class is as likely to contain stage IV metastatic tumors as it is to
contain those of stage I (**figure 3D, E**). This class likely represents a
group of tumors with aberrations in methylation, copy number alterations,
structural variants, or combinations thereof which we excluded from our
analysis. Consistent with this, kidney chromophobe and thyroid cancers have high
proportions of structural variations _vs._ other variant types [@PCAWG-MAIN] and
are heavily skewed to cluster 1 membership (**Supplementary figure S1**).
Likewise, Class 10 represents a group of hyper-mutated tumors that harbor so
many mutations that virtually no pathway is unaffected. It seems likely that a
significant fraction of the “mutant” samples for each pathway are burdened with
excess passenger mutations. This could be addressed with application of
ever-increasingly sophisticated filtering likely passenger mutations [ _e.g._
@spliceAI; @deep-learning-clinical-variants]. In the future, we hope to
incorporate these other data into a comprehensive pathway-centric analysis as we
have done here for point mutations and indels.

@@

By modeling CDC longevity data as a baseline risk function we showed that each
of our pathway disruption classes exhibit cancer-type specific effects on
expected survival (**figure 7**). Contrary to what one might expect at the
outset, we found that models assuming different survival based on class and
cancer type fit better than models in which they are independent factors.
However, considering that within each cancer type there are different clinical
standards of care, and even within classes of drugs the preferred treatment can
very between cancers, it makes sense that we observe tissue-specific class
effects. Contrast the situation with ovarian _vs._ breast cancer, which are both
hormonally driven cancers, for example. Ovarian cancer has but one main
treatment axis, platinum, whereas breast cancer patients have a variety of
treatment regimens based on molecular subtype and other factors. It is easy to
see how standard of care for one cancer and class could produce different
outcomes than the same class of a different cancer type. Unfortunately, given
the diversity of drug classes and treatments (including non-chemical treatments
such as surgery and palliative care), we lack sufficient power to explore these
variables in the TCGA data. It is our hope that future studies will help to
distinguish between treatment-specific effects on survival given different
pathway disruption classes.

>>>>

Finally, we must remark on the limitations of our work exemplified in two of the
classes, 1 and 10, for which we did not find many distinctive associations with
pathways (**figure 5B, C**). Class 1 had a relatively low proportion of mutated
pathways across the board, although it is broadly enriched in many of the same
tumor-promoting pathways common to the other groups. This likely does not
reflect a true difference in stages of tumor evolution, as our data show clearly
that this class is as likely to contain stage IV metastatic tumors as it is to
contain those of stage I (**figure 3D, E**). This class likely represents a
group of tumors with aberrations in methylation, copy number alterations,
structural variants, or combinations thereof which we excluded from our
analysis. Consistent with this, kidney chromophobe and thyroid cancers have high
proportions of structural variations _vs._ other variant types [@PCAWG-MAIN] and
are heavily skewed to cluster 1 membership (**Supplementary figure S1**).
Likewise, Class 10 represents a group of hyper-mutated tumors that harbor so
many mutations that virtually no pathway is unaffected. It seems likely that a
significant fraction of the “mutant” samples for each pathway are burdened with
excess passenger mutations. This could be addressed with application of
ever-increasingly sophisticated filtering of likely passenger mutations [ _e.g._
@spliceAI; @deep-learning-clinical-variants]. In the future, we hope to
incorporate these other data into a comprehensive pathway-centric analysis as we
have done here for point mutations and indels.

@@

By modeling CDC longevity data as a baseline risk function we showed that each
of our pathway disruption classes exhibit cancer-type specific effects on
expected survival (**figure 9**). Contrary to what one might expect at the
outset, we found that models assuming different survival based on class and
cancer type fit better than models in which they are independent factors.
However, considering that within each cancer type there are different clinical
standards of care, and even within classes of drugs the preferred treatment can
very between cancers, it makes sense that we observe tissue-specific class
effects. Contrast the situation with ovarian _vs._ breast cancer, which are both
hormonally driven cancers, for example. Ovarian cancer has but one main
treatment axis, platinum, whereas breast cancer patients have a variety of
treatment regimens based on molecular subtype and other factors. It is easy to
see how standard of care for one cancer and class could produce different
outcomes than the same class of a different cancer type. Unfortunately, given
the diversity of drug classes and treatments (including non-chemical treatments
such as surgery and palliative care), we lack sufficient power to explore these
variables in the TCGA data. It is our hope that future studies will help to
distinguish between treatment-specific effects on survival given different
pathway disruption classes.
